Lattice Boltzmann study of mass transfer for complicated geometries

### 2012-02-10 Text/paper.tex
Minor changes.
Started doing the Same Peclet simulations.

\newcommand{\beq}{\begin{equation}}
\newcommand{\feq}{\end{equation}}
\newcommand{\beqstar}{\begin{equation*}}
\newcommand{\feqstar}{\end{equation*}}                                     
\newcommand{\beqal}{\begin{equation}\begin{aligned}}
\newcommand{\feqal}{\end{aligned}\end{equation}}
\newcommand{\beqalstar}{\begin{equation*}\begin{aligned}}
\newcommand{\feqalstar}{\end{aligned}\end{equation*}}
\newcommand{\vol}{k_L\,a}
\newcommand{\lbubble}{L_{\mathrm{bubble}}}
\newcommand{\lunit}{L_{\mathrm{unit}}}
\newcommand{\lslug}{L_{\mathrm{slug}}}
\newcommand{\ububble}{U_{\mathrm{bubble}}}
\newcommand{\uliq}{U_{\mathrm{liq}}}
\newcommand{\ugas}{U_{\mathrm{gas}}}
\newcommand{\uoutlet}{U_{\mathrm{outlet}}}
\newcommand{\cbubble}{C_{\mathrm{bubble}}}
\newcommand{\cinlet}{C_{\mathrm{inlet}}}
\newcommand{\coutlet}{C_{\mathrm{outlet}}}
\newcommand{\coverall}{C_{\mathrm{overall}}}
\newcommand{\cstar}{C^{*}}
\newcommand{\cmedium}{C_{\mathrm{medium}}}
\newcommand{\volnondim}{\vol \frac{\lunit}{\ububble+\ugas}}
\newcommand{\omegaplus}{\omega_{+}}
\newcommand{\omegaminus}{\omega_{-}}
\newcommand{\holdup}{\varepsilon_{\mathrm{gas}}}

@@

\end{description}

Overall work goal is to compare numerical simulations with given correlations and to establish a
new one.

>>>>

\newcommand{\beq}{\begin{equation}}
\newcommand{\feq}{\end{equation}}
\newcommand{\beqstar}{\begin{equation*}}
\newcommand{\feqstar}{\end{equation*}}                                     
\newcommand{\beqal}{\begin{equation}\begin{aligned}}
\newcommand{\feqal}{\end{aligned}\end{equation}}
\newcommand{\beqalstar}{\begin{equation*}\begin{aligned}}
\newcommand{\feqalstar}{\end{aligned}\end{equation*}}
\newcommand{\vol}{k_L\,a}
\newcommand{\lbubble}{L_{\mathrm{bubble}}}
\newcommand{\lunit}{L_{\mathrm{unit}}}
\newcommand{\lslug}{L_{\mathrm{slug}}}
\newcommand{\lfilm}{L_{\mathrm{film}}}
\newcommand{\ububble}{U_{\mathrm{bubble}}}
\newcommand{\uliq}{U_{\mathrm{liq}}}
\newcommand{\ugas}{U_{\mathrm{gas}}}
\newcommand{\uoutlet}{U_{\mathrm{outlet}}}
\newcommand{\cbubble}{C_{\mathrm{bubble}}}
\newcommand{\cinlet}{C_{\mathrm{inlet}}}
\newcommand{\coutlet}{C_{\mathrm{outlet}}}
\newcommand{\coverall}{C_{\mathrm{overall}}}
\newcommand{\cstar}{C^{*}}
\newcommand{\cmedium}{C_{\mathrm{medium}}}
\newcommand{\volnondim}{\vol \frac{\lunit}{\ububble+\ugas}}
\newcommand{\omegaplus}{\omega_{+}}
\newcommand{\omegaminus}{\omega_{-}}
\newcommand{\holdup}{\varepsilon_{\mathrm{gas}}}
\newcommand{\volfilm}{k_L\, a_{\mathrm{film}}}
\newcommand{\volcap}{k_L\, a_{\mathrm{cap}}}
\newcommand{\tfilm}{t_{\mathrm{film}}}
\newcommand{\tcap}{t_{\mathrm{cap}}}

@@

One can see in Appendix that for the two-dimensional geometry one can estimate the impact to the volumetric mass transfer coefficient:
\beq
\vol=\volfilm+\volcap,
\feq
where $\volfilm$ can be calculated as in Appendix {\color{red} give appendix number here}. For
example, for the unsaturated film the expression for the volumetric mass transfer coefficient is as:
\beq
k_L=2\sqrt{\frac{D}{\pi \tfilm}},
\feq
where $\tfilm$ is the contact time of the bubble with the film:
\beq
\tfilm=\frac{\lfilm}{\ububble}=\frac{\lbubble-H(1-2\delta)}{\ububble}
\feq
Therefore the volumetric mass transfer coefficient can be calculated as (considering that there are
two films):
\beqal
\volfilm&=2\sqrt{\frac{D \ububble}{\pi (\lbubble-H(1-2\delta))}} 2 \frac{
\lbubble-H(1-2\delta)}{\lunit\,H}\\
&=4\sqrt{\frac{D \ububble}{\pi}} \frac{
\sqrt{\lbubble-H(1-2\delta)}}{\lunit\,H}
\feqal
The same expression for the coefficient $k_L$ can be used to calculate mass transfer from the caps:
\beq
k_L=2\sqrt{\frac{D}{\pi \tcap}},
\feq
where $\tcap$ is the time when the fluid interacts with bubble caps. It can be estimated due to
following explanations: if there is a vortex in front of the bubble then the fluid packet travels
the distance as half of the circumference and then is leaving with the bubble velocity from the
bubble surface. Therefore, $\tcap$ can be estimated as:
\beq
\tcap=\frac{\pi H(1-2\delta)}{2 \ububble}.
\feq
\end{description}
Therefore, the cap volumetric mass transfer coefficient is the following (considering that two rear
and front bubble caps form a whole circle):
\beqal
\volcap&=2 \sqrt{\frac{2 D \ububble}{\pi^2 H(1-2\delta) }} \frac{\pi H(1-2\delta)}{\lunit
H}=2\sqrt{2} \sqrt{D \ububble}\frac{\sqrt{H (1-2\delta)}}{\lunit H}
\feqal
Therefore, the whole volumetric mass transfer correlation is the sum of contributions from film and
width:
%\beq
\begin{multline}
\volnondim=\\
\frac{\lunit}{\ububble+\ugas}\Bigl(4\sqrt{\frac{D \ububble}{\pi}} \frac{
\sqrt{\lbubble-H(1-2\delta)}}{\lunit\,H}+2\sqrt{2} \sqrt{D \ububble}\frac{\sqrt{H
(1-2\delta)}}{\lunit H}\Bigr)
\end{multline}
%\feq

Some results:
200
Bubble velocities= [0.0041703652250000001, 0.0097256202740000006, 0.01588329539, 0.02169746379,
0.028008169950000002, 0.034069223429999997, 0.040563539609999998, 0.047033809750000002,
0.053112798260000003]
Liquid velocities= [0.003537420725238904, 0.007658332369944002, 0.011807274133101,
0.01596516025341501, 0.020125845424484007, 0.024289948286241009, 0.028452175441002994,
0.03260953169443101, 0.036767336801079013]
Gas velocities= [0.00045059406134383335, 0.0010532846756742, 0.0017299026452428667,
0.0023709541932125998, 0.0030704423110520002, 0.0037603337540471999, 0.0044949810359827995,
0.005218714750494167, 0.0059203065793813339]
Holdups= [0.10804666666666667, 0.10829999999999999, 0.10891333333333333, 0.10927333333333333,
0.10962666666666666, 0.11037333333333334, 0.11081333333333333, 0.11095666666666666,
0.11146666666666667]
Widths= [0.083691708460646039, 0.11453052579506764, 0.13337291846007834, 0.14494545818883292,
0.15412449600357994, 0.16018690601288671, 0.16531042826765582, 0.16967458296893251,
0.17344750155827895]
Capillaries= [0.073722338265551551, 0.1719263011747692, 0.28077964694645074, 0.38356059451146424,
0.49511917250675719, 0.60226447292781515, 0.71706884817900307, 0.83144814548157386,
0.9389104954359766]
Reynolds= [1.2511095675000001, 2.9176860822000004, 4.7649886170000002, 6.5092391369999998,
8.4024509850000015, 10.220767028999999, 12.169061883000001, 14.110142925, 15.933839478000003]
Bubble lengths= [2.1349999999999998, 2.2999999999999998, 2.4300000000000002, 2.5049999999999999,
2.5750000000000002, 2.6400000000000001, 2.6899999999999999, 2.7349999999999999, 2.7599999999999998]
Slug lengths= [12.865, 12.699999999999999, 12.57, 12.495000000000001, 12.425000000000001,
12.359999999999999, 12.31, 12.265000000000001, 12.24]
350
Bubble velocities= [0.0046341367029999999, 0.01080052902, 0.017377746239999999,
0.024331842780000001, 0.031265874229999997, 0.038061554249999997, 0.045243732449999997,
0.052329724549999998, 0.059326526419999998]
Liquid velocities= [0.0038803812007794009, 0.0084385532971600029, 0.013036803534905009,
0.017645755902851005, 0.02225813959945698, 0.026865244181491992, 0.031470323329384015,
0.03607200072948101, 0.040659019252402999]
Gas velocities= [0.00087688679406400325, 0.0020412639830166002, 0.0033112716202111999,
0.0046254022063354, 0.0060081546116175655, 0.0073038853887274994, 0.0087113790583645007,
0.010121615322460999, 0.011433012661486266]
Holdups= [0.18922333333333333, 0.18899666666666667, 0.19054666666666667, 0.19009666666666666,
0.19216333333333332, 0.19189666666666666, 0.19254333333333334, 0.19342000000000001,
0.19271333333333332]
Widths= [0.085334881242557656, 0.11754777714799797, 0.13562130389756313, 0.14771657697902627,
0.1559458293070895, 0.16216161772762192, 0.16690218748145516, 0.17071829922107651,
0.17395151438539647]
Capillaries= [0.081920737190919246, 0.19092818276110241, 0.30719805520107568, 0.4301302757125734,
0.55270779219396804, 0.67283957781686643, 0.79980375053962127, 0.92506757717322885,
1.0487547283956218]
Reynolds= [1.3902410109000001, 3.2401587060000003, 5.2133238720000001, 7.2995528340000009,
9.3797622690000004, 11.418466275, 13.573119734999999, 15.698917365, 17.797957926000002]
Bubble lengths= [3.625, 3.9100000000000001, 4.1200000000000001, 4.2649999999999997,
4.3899999999999997, 4.4749999999999996, 4.5549999999999997, 4.6200000000000001, 4.6550000000000002]
Slug lengths= [11.375, 11.09, 10.879999999999999, 10.734999999999999, 10.609999999999999, 10.525,
10.445, 10.379999999999999, 10.344999999999999]
500
Bubble velocities= [0.0052657317470000001, 0.01217878658, 0.01962143069, 0.027579066659999998,
0.03530242384, 0.042899137660000003, 0.050761276039999997, 0.064964573319999996,
0.080310501160000006]
Liquid velocities= [0.0043048060885056008, 0.0094048376047090026, 0.014562819556581999,
0.019728073069401001, 0.024892914265017, 0.030049067027628981, 0.035198125742582999,
0.044643609849748973, 0.054620418463494469]
Gas velocities= [0.0014211507887586733, 0.0033086109501886001, 0.0053329740520040667,
0.0075300964306241995, 0.0096365026356047996, 0.011760655592363467, 0.013922125975237332,
0.017439956256331733, 0.021556677019696671]
Holdups= [0.26988666666666666, 0.27167000000000002, 0.27179333333333333, 0.27303666666666665,
0.27296999999999999, 0.27414666666666665, 0.27426666666666666, 0.26845333333333332,
0.26841666666666669]
Widths= [0.089396777499460164, 0.12189600121336229, 0.14008376507935255, 0.15155603520007516,
0.15974866664944226, 0.16554914994206618, 0.17007608573781216, 0.30999461831205144,
0.22732300823418256]
Capillaries= [0.093085865655324646, 0.21529256443354303, 0.34686116743702095, 0.48753362635204572,
0.6240645822396409, 0.75835677866103002, 0.89734106273915526, 1.1484222582865664,
1.4197024992681524]
Reynolds= [1.5797195240999999, 3.6536359740000002, 5.8864292070000008, 8.2737199980000007,
10.590727152000001, 12.869741298000003, 15.228382812, 19.489371995999999, 24.093150348000005]
Bubble lengths= [5.1399999999999997, 5.5750000000000002, 5.8650000000000002, 6.085,
6.2400000000000002, 6.3550000000000004, 6.4500000000000002, 0.38500000000000001, 1.03]
Slug lengths= [9.8599999999999994, 9.4250000000000007, 9.1349999999999998, 8.9149999999999991,
8.7599999999999998, 8.6449999999999996, 8.5500000000000007, 14.615, 13.970000000000001]

650
Bubble velocities= [0.0058695922599999997, 0.014085166979999999, 0.022472028200000001,
0.031304453740000002, 0.04025350363, 0.049387996580000003, 0.1094018549, 0.15286911380000001,
0.099952447360000002]
Liquid velocities= [0.0048267119263748015, 0.010635775511508006, 0.016503260202420003,
0.022370510614027995, 0.028237291072309011, 0.034083362994406008, 0.046219448474894988,
0.067168537760430982, 0.067247061842290498]
Gas velocities= [0.0020611073220990002, 0.0049701389711193991, 0.0079327008613606677,
0.011103168000682334, 0.0142102918514626, 0.017489606602220133, 0.0380718455052,
0.053700371859376671, 0.035093637442920531]
Holdups= [0.35115000000000002, 0.35286333333333331, 0.35300333333333334, 0.35468333333333335,
0.35302, 0.35412666666666665, 0.34799999999999998, 0.35128333333333334, 0.35110333333333332]
Widths= [0.095686850123114092, 0.12704971860959374, 0.14492911134976988, 0.155909577069202,
0.16366636791645661, 0.16914779849407549, 0.1413349338842142, 0.18097719638547718,
0.25542130684882919]
Capillaries= [0.10376071224615181, 0.24899292714257107, 0.39725308818088317, 0.55338978802236449,
0.71158813458225767, 0.87306468227340051, 1.933969836854418, 2.702369675048951, 1.7669263331111857]
Reynolds= [1.7608776779999999, 4.2255500939999999, 6.7416084600000001, 9.391336122000002,
12.076051089000002, 14.816398974000002, 32.82055647, 45.860734140000005, 29.985734208]
Bubble lengths= [6.6900000000000004, 7.29, 7.6600000000000001, 7.9249999999999998,
8.1099999999999994, 8.2699999999999996, 0.53500000000000003, 1.645, 0.81000000000000005]
Slug lengths= [8.3099999999999987, 7.71, 7.3399999999999999, 7.0750000000000002, 6.8900000000000006,
6.7300000000000004, 14.465, 13.355, 14.19]

800
Bubble velocities= [0.0068458110670000004, 0.01639621631, 0.026605047409999999,
0.036840145929999998, 0.047125554479999998, 0.098226389829999997, 0.088886650849999999, 0.155738557,
0.1598356065]
Liquid velocities= [0.0054993290778350004, 0.012256252057892001, 0.019051401038488,
0.025834965312065014, 0.032605613362688031, 0.054062078686480969, 0.059582587094709517,
0.077607375795613484, 0.084522484007136495]
Gas velocities= [0.0029579152264591368, 0.0070805967053647672, 0.011559094948222701,
0.016054075992888966, 0.020507156287516797, 0.042450171471531663, 0.038447624536331333,
0.067548224763348333, 0.06939502609607251]
Holdups= [0.43207666666666666, 0.43184333333333336, 0.43447000000000002, 0.43577666666666665,
0.43515999999999999, 0.43216666666666664, 0.43254666666666669, 0.43372833333333333,
0.43416500000000002]
Widths= [0.10003982925931058, 0.13317387670353664, 0.15015330462158188, 0.16056724856591476,
0.16754248421583459, 0.17575649947662744, 0.091940069806373592, 0.18678182336855095,
0.10494658358534029]
Capillaries= [0.12101798570494038, 0.28984689346506176, 0.47031523593501479, 0.65124792517512475,
0.83306997849960196, 1.7364136585066581, 1.5713088393249, 2.7530947436726914, 2.8255210307803655]
Reynolds= [2.0537433201000002, 4.9188648930000003, 7.9815142230000005, 11.052043779,
14.137666343999999, 29.467916948999999, 26.665995254999999, 46.721567100000001, 47.950681950000003]
Bubble lengths= [8.2899999999999991, 9.0649999999999995, 9.5299999999999994, 9.8399999999999999,
10.055, 7.1200000000000001, 5.9649999999999999, 3.585, 5.4699999999999998]
Slug lengths= [6.7100000000000009, 5.9350000000000005, 5.4700000000000006, 5.1600000000000001,
4.9450000000000003, 7.8799999999999999, 9.0350000000000001, 11.414999999999999, 9.5300000000000011]

950
Bubble velocities= [0.0081105889820000006, 0.01968896174, 0.031648192290000002,
0.043746032820000003, 0.056154315660000002, 0.09110329466, 0.12485531280000001, 0.13740875180000001,
0.20420121150000001]
Liquid velocities= [0.0064009034784269989, 0.014478691581925994, 0.022535789859960999,
0.030564372867634991, 0.038537555847714004, 0.053671674539171009, 0.072182391854489547,
0.079464538286806011, 0.11495232139103347]
Gas velocities= [0.0041692753014937068, 0.0101189449966556, 0.016351882498422902,
0.022612178544548601, 0.029043012059352002, 0.046834078041163928, 0.064378728754688003,
0.07078199622721601, 0.10522080026172]
Holdups= [0.51405333333333336, 0.51393999999999995, 0.51667666666666667, 0.51689666666666667,
0.51719999999999999, 0.51407666666666663, 0.51562666666666668, 0.51512000000000002,
0.51527999999999996]
Widths= [0.10534960942798749, 0.13968846394555789, 0.15578372184740735, 0.16543619956323713,
0.17153030593666954, 0.19424048762229987, 0.19865767560879785, 0.17950947813380666,
0.12066555457759411]
Capillaries= [0.14337631171472745, 0.34805495902191214, 0.55946628451387026, 0.77332791142578172,
0.99267743490189853, 1.6104939360630546, 2.207150958701189, 2.4290665048039806, 3.6098015344539602]
Reynolds= [2.4331766946000006, 5.9066885220000005, 9.4944576870000024, 13.123809846000002,
16.846294698000001, 27.330988397999999, 37.456593840000004, 41.222625540000003, 61.26036345]
Bubble lengths= [9.9499999999999993, 10.914999999999999, 11.449999999999999, 11.789999999999999,
12.035, 8.9499999999999993, 9.1750000000000007, 6.2800000000000002, 4.6050000000000004]
Slug lengths= [5.0500000000000007, 4.0850000000000009, 3.5500000000000007, 3.2100000000000009,
2.9649999999999999, 6.0500000000000007, 5.8249999999999993, 8.7199999999999989, 10.395]

1100
Bubble velocities= [0.0097824175900000001, 0.024498111669999999, 0.039400354679999999,
0.054303837889999998, 0.09376646777, 0.11965875519999999, 0.1483564782, 0.17728152890000001,
0.225203454]
Liquid velocities= [0.0077186053895769999, 0.01770605859890001, 0.027603204679605003,
0.037367546771407988, 0.055859605054054476, 0.071491609706563986, 0.087203948180378035,
0.099228457857510055, 0.12788810850038299]
Gas velocities= [0.005819331967880567, 0.014661466551517132, 0.023538559892925599,
0.032476410250114496, 0.05588731523006054, 0.071375450320506656, 0.088544833188091004,
0.10585273262168618, 0.13442619372714001]
Holdups= [0.59487666666666672, 0.5984733333333333, 0.59741999999999995, 0.59804999999999997,
0.5960266666666667, 0.59649166666666664, 0.59683833333333336, 0.59708833333333333,
0.59691000000000005]
Widths= [0.11146768101104691, 0.14705522853564731, 0.16189749306607318, 0.17031901280973424,
0.22743073870670458, 0.14494577386922328, 0.15532348617470038, 0.19125412329860453,
0.089789110604593075]
Capillaries= [0.1729303453571891, 0.43306952220305733, 0.69650644938457806, 0.95996530041184946,
1.6575726302019211, 2.1152879307565264, 2.6225967942043553, 3.1339242816077229, 3.9810722367508178]
Reynolds= [2.9347252770000001, 7.349433501, 11.820106404000001, 16.291151367000001, 28.129940331,
35.897626559999999, 44.506943460000002, 53.184458670000005, 67.561036200000004]
Bubble lengths= [11.699999999999999, 12.875, 13.465, 13.82, 12.51, 11.305, 10.44,
9.6199999999999992, 14.234999999999999]
Slug lengths= [3.3000000000000007, 2.125, 1.5350000000000001, 1.1799999999999997,
2.4900000000000002, 3.6950000000000003, 4.5600000000000005, 5.3800000000000008, 0.76500000000000057]
